verisr
========================================================

This package is to support data analysis within the VERIS framework (http://veriscommunity.net).  It is intended to work directly with raw JSON and can be used against the VERIS Community Database (VCDB) found at (http://veriscommunity.net/doku.php?id=public) and (https://github.com/vz-risk/VCDB).

Install it from straight from github:
```{r echo=FALSE}
savetime <- proc.time()
```

```{r eval=FALSE}
library("devtools")
install_github("verisr", "jayjacobs")
```

To begin, load the package and point it at a directory of JSON files storing VERIS data.

```{r}
library(verisr)
vcdb.dir <- "../VCDB/data/json/"
vcdb <- json2veris(vcdb.dir)
```

Now that we have this, we can get a quick view of what's in the data:

```{r}
summary(vcdb)
```

And let's look at a high level bar plot of the data:

```{r "basic-plot"}
plot(vcdb)
```

Let's look for a specific variable:

```{r}
ext.variety <- getenum(vcdb, "actor.external.variety")
print(ext.variety)
```

And we could create a barplot with ggplot:

```{r "basic-ggplot"} 
library(ggplot2)
gg <- ggplot(ext.variety, aes(x=enum, y=x))
gg <- gg + geom_bar(stat="identity", fill="steelblue")
gg <- gg + coord_flip() + theme_bw()
print(gg)
```

In progress: searching by two enumerations:

```{r}
hacking.actor <- getenumby(vcdb, "action.hacking.vector", "actor.external.variety")
head(hacking.actor)
```

Now we can create all sorts of views of this data.
For example a faceted bar plot comparing the two:

```{r "facet-ggplot", fig.height=10}
gg <- ggplot(hacking.actor, aes(x=enum, y=x))
gg <- gg + geom_bar(stat="identity", fill="steelblue")
gg <- gg + facet_wrap( ~ primary, ncol=2)
gg <- gg + ylab("Incidents") + xlab("Hacking Vector") 
gg <- gg + ggtitle("External Actors by Hacking Vector")
gg <- gg + coord_flip() + theme_bw()
print(gg)
```

Finally, let's set a filter for only confirmed loss events (data_disclosure="Yes").
Then get the interaction of the top level actions and assets.

```{r}
ddfilter <- getfilter(vcdb, and=list("attribute.confidentiality.data_disclosure" = "Yes"))
action.asset <- getenumby(vcdb, enum="asset.assets", primary="action", filter=ddfilter)
head(action.asset)
```

```{r echo=F}
assetmap <- c("Server", "Network", "User Dev", "Media", 
              "Person", "Kiosk/Term", "Unknown")
action.asset$enum <- factor(action.asset$enum, 
                            levels=rev(assetmap), ordered=T)
action.asset$primary <- factor(action.asset$primary,
                               levels=c("malware", "hacking", "social", 
                                        "misuse", "physical", "error", 
                                        "environmental"),
                               ordered=T)
```

And now make a nice 2 x 2 grid with the data.

```{r  "a2grid", echo=FALSE}
slim.aa <- action.asset[which(action.asset$x>0), ]
gg <- ggplot(action.asset, aes(x = primary, y = enum, fill = x, label = x))
gg <- gg + geom_tile(fill="white", color="gray80")
gg <- gg + geom_tile(data=slim.aa, color="gray80") + geom_text(data=slim.aa)
gg <- gg + scale_fill_gradient(low = "#D8EEFE", high = "#4682B4")
gg <- gg + scale_x_discrete(expand=c(0,0)) + scale_y_discrete(expand=c(0,0))
gg <- gg + ylab("") + xlab("")
gg <- gg + ggtitle("A2 Grid: Assets and Actions (confirmed data loss)")
gg <- gg + theme_bw()
gg <- gg + theme(axis.text.x = element_text(angle = 90, hjust = 1, vjust=0.5),
                 axis.text = element_text(size=14),
                 axis.ticks = element_blank(),
                 legend.position = "none",
                 plot.background = element_blank(),
                 panel.grid.major = element_blank(),
                 panel.grid.minor = element_blank(),
                 panel.border = element_rect(colour = "gray30"),
                 panel.background = element_blank())
print(gg)
```

```{r echo=FALSE}
print(proc.time() - savetime)
```
